---
title: Modelling the sex-age distribution of the global refugee and asylum-seeker
  population
author: "Sebastian Steinmueller (UNHCR)"
date: 'Version: `r Sys.Date()`'
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

# Introduction




# Data set
- Describe UNHCR official population statistics content and compilation
- 

## Descriptives of age distribution and missingness


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
## run code to create descriptive figures and tables

source("descriptives.R")

```




Globally
- % per age/sex cat. 
- % age missing
- % age and sex missing

```{r, echo=F}

knitr::kable(t.checkunknowns.ref, caption = 'Unknown age in disaggregated data')

```

In some reporting countries, the age for a small proportion of refugees for whom age-disaggregated data was generally available could not be determined. This was the case for `r round(t.checkunknowns.ref$freq.ageUnknown*100,1)` per cent of the `r as.numeric(t.typeOfDisaggregation[t.typeOfDisaggregation$typeOfDisaggregation=="Sex/Age"&!(is.na(t.typeOfDisaggregation$typeOfDisaggregation)),])` refugees with available age disaggregation.

By CoO
- % per age/sex cat. 
- % age missing
- % age and sex missing

```{r, fig.height = 7, fig.width = 15, fig.align='center', echo=F, fig.cap='Demographic disaggregation coverage by origin country and asylum region'}
p.typeOfDisaggregation.ori.asyreg

```



```{r, fig.height = 7, fig.width = 15, echo=F, fig.cap='2020 end-year refugee/Venezuelan population by origin country and asylum region'}
p.totalEndYear.ori.asyreg

```

By CoA
- % per age/sex cat. 
- % age missing
- % age and sex missing


Discuss types and reasons for missingness (NMAR) and outline modelling approach to overcome.
Why is using available data so bad?
