---
title: Modelling the sex-age distribution of the global refugee and asylum-seeker
  population
author: "Sebastian Steinmueller (UNHCR)"
date: 'Version: `r Sys.Date()`'
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
  github_document:
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
## run code to create descriptive figures and tables

source("descriptives.R")

```


# Outline of descriptive analysis
- Describe UNHCR official population statistics content and compilation
- Show and explore structure of data sets: especially asylum - origin - location. 
- Show origin and asylum country structure in missing and non-missing demographic data
- Explain choice of age brackets (18-59 or finer?): show population and country coverage



# Introduction




# Data set
UNHCR compiles official statistics on stocks and flows of forcibly displaced and stateless persons twice a year, once for mid-year figures (Mid-Year Statistical Reporting, MYSR) and once for end-year figures (Annual Statistical Reporting, ASR). For these reporting exercises, country operations compile aggregate population figures from a range of sources and data producers such as governments, UNHCR's own refugee registration database proGres and sometimes non-governmental actors. The figures undergo a statistical quality control process at the country, regional and global level of the organisation and are disseminated on the publicly available refugee data finder (https://www.unhcr.org/refugee-statistics/) after undergoing a light statistical disclosure control process to suppress very small counts of persons that could identify individuals. 

The end-year figures compiled with reporting date 31 December contain sex- and age breakdowns of the stocks of displaced and stateless people under UNHCR's mandate. Table \@ref(tab:demref2020)) displays the sex- and age-disaggregated data on the stock of refugees under UNHCR's mandate (including Venezuelans displaced abroad, excluding Palestine refugees under UNRWA's mandate). The data is available on sub-national level as indicated by the variables location and urbanRural. Variable statelessStatus displays whether the reported population is stateless (STL and UDN) or not stateless (NSL). The variables [sex]_[agebracket] contain the counts of refugees as of 31 December 2020 in the individual sex and age brackets in the respective geographic/stateless combination. For example, female_12_17 contains the number of female refugees aged 12 to 17. Variable totalEndYear is the total number of refugees over all sex/age categories.   
 
```{r demref2020, cols.min.print=32, rows.print=7, rownames.print = F, caption = "Demographic breakdown of UNHCR refugee population end-2020"}
demref2020 %>% 
  select(asylum_country, origin_country, location, urbanRural, statelessStatus, female_0_4:totalEndYear, typeOfDisaggregation, typeOfDisaggregationBroad)
``` 
 

# Geographic distribution of missing data

## By region of asylum

```{r, fig.height = 7, fig.width = 15, fig.align='center', echo=F, fig.cap="Demographic disaggregation coverage by region of asylum"}
p.typeOfDissaggregationBroad.asyregion

```





Globally
- % per age/sex cat. 
- % age missing
- % age and sex missing


## By region of origin

```{r, fig.height = 7, fig.width = 15, fig.align='center', echo=F, fig.cap='Demographic disaggregation coverage by region of origin'}
p.typeOfDissaggregationBroad.oriregion

```



By CoO
- % per age/sex cat. 
- % age missing
- % age and sex missing

```{r, fig.height = 7, fig.width = 15, fig.align='center', echo=F, fig.cap='Demographic disaggregation coverage by origin country and asylum region'}
p.typeOfDisaggregationBroad.ori.asyreg

```



```{r, fig.height = 7, fig.width = 15, echo=F, fig.cap='2020 end-year refugee/Venezuelan population by origin country and asylum region'}
p.totalEndYear.ori.asyreg

```

By CoA
- % per age/sex cat. 
- % age missing
- % age and sex missing


Discuss types and reasons for missingness (NMAR) and outline modelling approach to overcome.
Why is using available data so bad?
